Refinement for open automata

### 2021-06-16 2021-Refinement/report.tex
a bit of organisation in sec 6 + a few comments

Intuitively the right automaton is a refinement of the left automaton because the left automaton is what happens when there was no bound \(n\) in the right automaton.
\[ R \defobject \mbrc{\mpar{l, ||} \mapsto \top, \mpar{r, ||} \mapsto \top} \]
\(R\) is a hole-identical simulation of \(n\) control switches by parallel composition.
There are enough examples of proofs of refinement in this article for the reader to be able to prove it as an exercise.
\end{exi}
\begin{exi}[Example of hole-subset simulation]
Example of Appendix \ref{apx:composition} can be reused here.
The automaton on Figure \ref{fig:tlf} is a refinement of the one on Figure \ref{fig:tls} by limited composition correctness (Proposition \ref{prop:cc'}).

@@

The automaton on the bottom is a refinement of the automaton on the top using hole-superset simulation.
The relation on states is the same as the one for reflexivity.
Intuitively the automaton on the bottom binds some variables to action produced by the holes, the automaton is no more free to take a transition some hole is not able to perform.
For instace if the automaton in the hole \(mon\) can only perform \(\act{credit}\mpar{100}\) then after the composition, the automaton on the bottom should be a refinement of the one on the top.
Actually after such a composition it is still in hole-superset simulation with the top automaton by Proposition \ref{prop:cc'}.
\end{exi}
\begin{thm} A hole-\(\triangle\) simulation is a pre-simulation. \end{thm}
Proof is given in Appendix \ref{apx:presim}.
\begin{thm} The hole-\(\triangle\) simulation on automata is a preorder. \end{thm}
The idea for reflexivity is that the related automata are always in the same configuration.
\[ R \defobject \mpar{s, s'} \mapsto \choice{\bigwedge_{v \in V} v_1 = v_2 & \text{if } s = s'\\ \bot & \text{otherwise}} \]
Where \(v_x\) is the renaming to distinguish conflicting automaton variables.
\(R\) is a witness of the reflexivity, the proof is mainly the same as several of the following ones so it will be omitted.

Let \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\).
Let \(R_{12}\) be a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier on states as we did in deadlock prevention, which is possible because states are finite.
\[ R \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \]
\(R\) is a witness of the transitivity, proof is given in Appendix \ref{apx:trans}.
\begin{prop}[Limited composition correctness]\label{prop:cc'} Hole-subset simulation is correct w.r.t.\@ composition with automata without holes. \end{prop}
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\) with \(J_2 = \emptyset\), let \(k \in J_1\) and \(A \defobject A_1\subst{A_2}{k}\).
Let \(\reach{A}\) be a witness that \(A\) is non-locking.
The idea for the relation on configurations is that the copy of \(A_1\) in \(A\) do the same things as \(A_1\), as if we were trying to prove reflexivity.
\[ R \defobject \mpar{\mpar{s_{1'}, s_2}, s_1} \mapsto \choice{\reach{A}\mpar{s_{1'}, s_2} \wedge \bigwedge_\subbox{v_1 \in V_1} v_{1'} = v_1 & \text{if } s_{1'} = s_1 \\ \bot & \text{otherwise}} \]
Where \(y_{1'}\) is the renaming of \(y_1\) from \(A_1\) in \(A\).
\(R\) is a witness of the limited composition correctness, proof is given in Appendix \ref{apx:lchs}.
\begin{prop}[Special composition correctness]
For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3\), let \(k \in J_1 \setminus J_2\) and \(R_{12}\) be a hole-superset simulation of \(A_1\) by \(A_2\).
There is a hole-superset simulation of \(A_1\subst{A_3}{k}\) by \(A_2\).
\end{prop}
Let \(\reach{A_1\subst{A_3}{k}}\) be a witness that \(A_1\subst{A_3}{k}\) is non-locking.
The idea for the relation on configurations is to combine the non-locking composition and the already existing realtion on configurations.
\[ R \defobject \mpar{\mpar{s_1, s_3}, s_2} \mapsto R_{12}\mpar{s_1, s_2} \wedge \reach{A_1\subst{A_3}{k}}\mpar{s_1, s_3} \]
\(R\) is a witness of the secial composition correctness, proof is given in Appendix \ref{apx:schs}.

For composition correctness, the hole-subset simulation can only tell if there is a refinement when the composed automata have no holes.
To lift this limitation we need to allow adding holes which doesn't have a direct interpretation.
At first a solution using a function mapping the action performed by the new holes in the composition to the ones performed by removed hole in the base automaton was explored.
It was abandonned because the base automaton has nothing intelligent to tell about the action required by the composed automata.
That is the reason the hole-superset simulation doesn't constrain the action performed by the added holes.

>>>>

Intuitively the right automaton is a refinement of the left automaton because the left automaton is what happens when there was no bound \(n\) in the right automaton.
\[ R \defobject \mbrc{\mpar{l, ||} \mapsto \top, \mpar{r, ||} \mapsto \top} \]
\(R\) is a hole-identical simulation of \(n\) control switches by parallel composition.
There are enough examples of proofs of refinement in this article for the reader to be able to prove it as an exercise.
\end{exi}
\begin{exi}[Example of hole-subset simulation]
Example of Appendix \ref{apx:composition} can be reused here.
The automaton on Figure \ref{fig:tlf} is a refinement of the one on Figure \ref{fig:tls} by limited composition correctness (Proposition \ref{prop:cc'} below).

@@

The automaton on the bottom is a refinement of the automaton on the top using hole-superset simulation.
The relation on states is the same as the one for reflexivity.
Intuitively the automaton on the bottom binds some variables to action produced by the holes, the automaton is no more free to take a transition some hole is not able to perform.
For instace if the automaton in the hole \(mon\) can only perform \(\act{credit}\mpar{100}\) then after the composition, the automaton on the bottom should be a refinement of the one on the top.
Actually after such a composition it is still in hole-superset simulation with the top automaton by Proposition \ref{prop:cc'}.
\end{exi}
\begin{thm} A hole-\(\triangle\) simulation is a pre-simulation. \end{thm}
Proof is given in Appendix \ref{apx:presim}.
\begin{thm} The hole-\(\triangle\) simulation on automata is a preorder. \end{thm}

\paragraph{Proof sketch}
The idea for reflexivity is that the related automata are always in the same configuration.
\[ R \defobject \mpar{s, s'} \mapsto \choice{\bigwedge_{v \in V} v_1 = v_2 & \text{if } s = s'\\ \bot & \text{otherwise}} \]
Where \(v_x\) is the renaming to distinguish conflicting automaton variables.
\(R\) is a witness of the reflexivity, the proof is mainly the same as several of the following ones so it will be omitted.

For transitivity we introduce the following.
Let \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\).
Let \(R_{12}\) be a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier on states as we did in deadlock prevention, which is possible because states are finite.
\[ R \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \]
\(R\) is a witness of the transitivity, proof is given in Appendix \ref{apx:trans}.
\begin{prop}[Limited composition correctness]\label{prop:cc'} Hole-subset simulation is correct w.r.t.\@ composition with automata without holes. \end{prop}
\paragraph{Proof sketch}
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\) with \(J_2 = \emptyset\), let \(k \in J_1\) and \(A \defobject A_1\subst{A_2}{k}\).
Let \(\reach{A}\) be a witness that \(A\) is non-locking.
The idea for the relation on configurations is that the copy of \(A_1\) in \(A\) do the same things as \(A_1\), as if we were trying to prove reflexivity.
\[ R \defobject \mpar{\mpar{s_{1'}, s_2}, s_1} \mapsto \choice{\reach{A}\mpar{s_{1'}, s_2} \wedge \bigwedge_\subbox{v_1 \in V_1} v_{1'} = v_1 & \text{if } s_{1'} = s_1 \\ \bot & \text{otherwise}} \]
Where \(y_{1'}\) is the renaming of \(y_1\) from \(A_1\) in \(A\).
\(R\) is a witness of the limited composition correctness, proof is given in Appendix \ref{apx:lchs}.
\begin{prop}[Special composition correctness]
For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3\), let \(k \in J_1 \setminus J_2\) and \(R_{12}\) be a hole-superset simulation of \(A_1\) by \(A_2\).
There is a hole-superset simulation of \(A_1\subst{A_3}{k}\) by \(A_2\).
\end{prop}
\paragraph{Proof sketch}
Let \(\reach{A_1\subst{A_3}{k}}\) be a witness that \(A_1\subst{A_3}{k}\) is non-locking.
The idea for the relation on configurations is to combine the non-locking composition and the already existing realtion on configurations.
\[ R \defobject \mpar{\mpar{s_1, s_3}, s_2} \mapsto R_{12}\mpar{s_1, s_2} \wedge \reach{A_1\subst{A_3}{k}}\mpar{s_1, s_3} \]
\(R\) is a witness of the secial composition correctness, proof is given in Appendix \ref{apx:schs}.

\smallskip
For composition correctness, the hole-subset simulation can only tell if there is a refinement when the composed automata have no holes.
To lift this limitation we need to allow adding holes which doesn't have a direct interpretation.
At first a solution using a function mapping the action performed by the new holes in the composition to the ones performed by removed hole in the base automaton was explored.
It was abandonned because the base automaton has nothing intelligent to tell about the action required by the composed automata.
That is the reason the hole-superset simulation doesn't constrain the action performed by the added holes.
